---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r initialize the playground,echo = FALSE}
packages.used=c("data.table", 
                "dplyr", 
                "plyr",
                "tidytext",
                "tm",
                "readxl",
                "wordcloud",
                "lubridate")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# load packages
library(data.table)
library(dplyr)
library(plyr)
library(tidytext)
library(tm)
library(readxl)
library(wordcloud)
library(lubridate)
```

+ this file needs for the analysis of the diff parties and different years
+ the key point is to create the functions to subset the dataset automatically by the factor I choose
+ then I can just apply the functions I wrote in wordcloud file

+ 01192018 Ming Li

```{r}
load("data/20180119.RData")
```

So here I will use the dates and general.info file

Combine them with the original text file, we may find something

```{r general.info}
unique.party <- unique(general.info$Party)
unique.party
summary(as.numeric(general.info$Words))
unique.president <- unique(general.info$President)
length(unique.president)
```
So from the general.info file, we can see that there are 6 values for party and 40 individuals for president.

The maximum of speech words is 8460 and the minmum is 135. So we can create two factors to analyse: the party to whom the presidents belong, and the words of their speech.

```{r dates}
head(dates)

attr.name <- colnames(dates)[-1]
id.vars <- colnames(dates)[1]
melt.dates <- melt(dates, variable.name = id.vars, measure.vars = attr.name)
melt.dates <- data.table(melt.dates)
melt.dates$value <- mapvalues(melt.dates$value, from = "", to = NA)
melt.dates <- melt.dates[which(is.na(melt.dates$value) ==  F)]

```

